### 2019-02-17 STAT0030_ICA2/STAT0030_ICA2.Rmd
STAT0030_begin

---
title: "STAT0030_ICA2"
author: "Hongwei Peng"
date: "10/02/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r table}
library(knitr)
kable(cars)
```

```{r table2}
kable(PlantGrowth, col.names=c("Weight", "Group"), align=c("c","c"), 
      caption = "Table 2. Plant Growth Data")#this is a test
```