### 2019-02-18 STAT0030_ICA2/STAT0030_ICA2.Rmd
Hongwei
vision  2 question 1 c

---
title: "STAT0030_ICA2"
author: "Hongwei Peng"
date: "10/02/2019"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

# R Question 1

## Question 1a read the data
```{R question_1_a}
setwd('/Users/hongwei/Documents/GitHub/STAT/STAT0030_ICA2')
rawdata <- read.table("cars.dat", #input data
                      header=TRUE) #the first line as the names of the variables 
```

##  Question 1b

```{R question_1_b}
summary(rawdata)
table(rawdata$tr)
plot(rawdata,main="Plot Between all Variables") #overlook
pairs(rawdata[,2:4], # plot hp, wt, mpg 
      main = "Plot Between all Variables -- 3 species", #add the main title
      pch = c(21,24)[unclass(rawdata$tr)+1], #different tr shows different shape
      bg = c("red", "green3")[unclass(rawdata$tr)+1]) #different tr shows different colour
logdata <- cbind(rawdata[,1],log(rawdata[,c(2,3,4)])) #log the data
names(logdata) <- c("tr","hp","wt","mpg")#rename the names of the variables 
plot(logdata,main="Plot Between log Variables") #overlook
pairs(logdata[,2:4], # plot log(hp), log(wt), log(mpg)
      main = "Plot Between log Variables -- 3 species", #add the main title
      pch = c(21,24)[unclass(logdata$tr)+1], #different tr shows different shape
      bg = c("red", "green3")[unclass(logdata$tr)+1]) #different tr shows different colour
```

##  question 1c

```{R question_1_c}
boxplot(mpg~tr, #MPG by TR
        data=rawdata, #set the dataset
        xlab="Gear Transmission Type" , #add the xlab title
        ylab="Miles Per Gallon", #add the ylab title
        main="MPG by Gear Transmission Type", #add the main title
        names=c("Automatic","Manual")) #change xlab value to character
t.test(mpg~tr, data=logdata)
model<-lm(mpg~tr+hp+wt, data=logdata); # i.e, full without qsec and gears
summary(model)
summary(model$residuals)
par(mfrow=c(2,2)) #put 4 graphes together
plot(model) #plot 4 graphes as following

best<-step(model, direction="both")
summary(best)
summary(best$residuals)
par(mfrow=c(2,2)) #put 4 graphes together
plot(best)#plot 4 graphes as following

```