### 2019-03-22 STAT0017_ICA1/SN17052480_NEW.Rmd

---
title: "SN17052480"
author: "Hongwei Peng"
date: "21/03/2019"
output:
  pdf_document: default
  html_document: default
---

# Part A
## Question 1 (a)
From the question we can see:
$$ F(x)= P(X \leq x)= \begin{cases} 1-e^{1/x}, & \text {for $x < 0$ .} \\ 1, & \text{for $x \geq 0$ .} \end{cases} \tag {1}$$
$$ M_n=max \lbrace x_1,...,x_n \rbrace $$ 
Hazard Function:
\begin{align}
h(x) & =\frac{1-F(x)}{f(x)} \\
  & = \frac{e^{1/x}}{e^{1/x} \times (-x^{-2})} \\
  & =-x^{2}
\end{align} 
Derivative of h(x):
$h^{'}= -2x$

\begin{align}
\lim_{x \to x^F} h^{'}(x) \to \xi \\
Let\ \xi = -2M_n \\
We\ can\ see:\  x \sim GEV(0,1,-2M_n)
\end{align}

From the slides we have:
$$\begin{cases} 1-F(b_n)=1/n \\ a_n=h(b_n) \end{cases}$$
solve it we have: 
$$\begin{cases} b_n=- \frac{1}{log(n)} \\ a_n=- \frac{1}{{log^2(n)}} \end{cases}$$

## Question 1 (b)
When $\xi < 0$ the GEV distribution has light upper tail with the finite upper limit which is  $\mu - \sigma / \xi$.

Let $\mu - \sigma / \xi = 0$ then this distribution with a finite upper end point of 0  has an upper end point of infinity.

## Question 1 (c)
We assume $A = \{ x: 0< F(x)<1 \}$ , and $x^*=sup_{x \in A} A$.

Here F(x) can be any function including the question's function(1).

For $\forall x, x<x^*$, we have $Pr(M_n \leq x)=F^n(x) \to 0, as \  n \to \infty$.

For $\forall x, x \geq x^*$, we have $Pr(M_n \leq x)=F^n(x) \to 1 , as \ n \to \infty$.

This is mean whatever x or F(x), $M_n = 0 or 1$ when $n \to \infty$.

The $M_n$ is Degenerate distribution, which is useless.

That is why we need fit a GEV model.

## Question 1 (d)
The log-likelihood:
$$l( \phi ) = l(\mu , \sigma, \xi)=log L(\mu , \sigma, \xi)=log \prod_{i=1}^m g(z_i;\mu , \sigma, \xi)$$

# Part B
## Task 1
```{r}
setwd("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1")
raw_wm <- load("wm.Rdata")
raw_pot <- load("pot.Rdata")
```

## Task 2

## Task 3.1

## Task 3.2

## Task 3.3

## Task 4.1

## Task 4.2