### 2019-03-22 STAT0017_ICA1/SN17052480.Rmd
finished the exploratory

---
output:
  pdf_document: default
  html_document: default
---

<!-- saved from url=(0100)https://moodle-1819.ucl.ac.uk/pluginfile.php/1342515/mod_resource/content/7/stat0017ica1template.Rmd -->
<html><head><meta http-equiv="Content-Type" content="text/html; charset=UTF-8"></head><body>---
title: "STAT0017 ICA 1 2018-19"
author: "Student number: xxxxxxxx"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_width: 6
    fig_height: 4
---

```{r, include = FALSE}
# Load the data (you will need to edit this to point to the correct directory
# on your computer)
setwd("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1")
load("wm.Rdata")
load("pot.Rdata")
# You will also need to load any R libraries that you want to use.
library("evd")
library("evdbayes")
library("ismev")
library("revdbayes")
library("threshr")
library("bayesplot")
library("coda")
source("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1/pjnfunctions.R")
# Some parts of the analyses are based on simulation.
# We set a set a random number seed so that
# 1. The numeric results on which you comments will not change each time you run
#    your code (unless you change the order of the analyses)
# 2. Your work is easier for me to mark.
#### PLEASE DO NOT CHANGE THE FOLLOWING LINE OF CODE (KEEP THE SEED as 17) ####
set.seed(17)
```

---
# [This is a comment block that will not appear in your PDF file]
# Add your text (perhaps with LaTeX code in it) and R code in the 
# relevant sections below.
# There is an example of an R code `chunk' in the Winter maxima (wm) section
---

# Extremal Types Example

${}$\hspace*{\fill} [25]

---
# You could put LaTeX code into this section to give your answer to Part A.
# You can only use basic LaTeX code (no extra packages).
# If you don't do this then please don't delete the Extremal Types Example
# heading above: just leave this section blank.
---

# Exploratory analysis

## Winter maxima (wm)

```{r}
# Add R code here (and similarly elsewhere)
plot(wm$waterYear,wm$Hs,bty="l",pch=20,xlab="year",ylab="wave height (in metres)")
#(wm$Hs~wm$meanNAO) 
plot(wm$meanNAO,wm$Hs,bty="l",pch=20,xlab="year",ylab="the mean of NAO")
abline(lm(wm$Hs~wm$meanNAO), col="red") # regression line (wm$Hs~wm$meanNAO) 
lines(lowess(wm$Hs~wm$meanNAO), col="blue") # lowess line (wm$Hs~wm$meanNAO)
#(wm$Hs~wm$maxNAO)
plot(wm$maxNAO,wm$Hs,bty="l",pch=20,xlab="year",ylab="the max of NAO")
abline(lm(wm$Hs~wm$maxNAO), col="red") # regression line (wm$Hs~wm$maxNAO)
lines(lowess(wm$Hs~wm$maxNAO), col="blue") # lowess line (wm$Hs~wm$maxNAO)
#(wm$Hs~wm$meanA0)
plot(wm$meanA0,wm$Hs,bty="l",pch=20,xlab="year",ylab="the mean of AO")
abline(lm(wm$Hs~wm$meanA0), col="red") # regression line (wm$Hs~wm$meanA0)
lines(lowess(wm$Hs~wm$meanA0), col="blue") # lowess line (wm$Hs~wm$meanA0)
#(wm$Hs~wm$maxAO)
plot(wm$maxAO,wm$Hs,bty="l",pch=20,xlab="year",ylab="the max of AO")
abline(lm(wm$Hs~wm$maxAO), col="red") # regression line (wm$Hs~wm$maxAO)
lines(lowess(wm$Hs~wm$maxAO), col="blue") # lowess line (wm$Hs~wm$maxAO)

```

## Storm peaks (pot)

## Comments

${}$\hspace*{\fill} [10]

# Extreme value (EV) modelling of $H_s$

## GEV modelling of winter maxima

### Maximum Likelihood-Based Inference

### Comments

### Bayesian Inference

### Comments

## Binomial-GP modelling of storm peaks

### Threshold selection

### Maximum Likelihood-Based Inference

### Comments

### Bayesian Inference

### Comments

${}$\hspace*{\fill} [25]

## Reporting to your client

${}$\hspace*{\fill} [15]

# EV regression modelling of winter maximum $H_s$ on NAO

## Build a GEV regression model

${}$\hspace*{\fill} [15]

## Inference for $H_s^{100}$

${}$\hspace*{\fill} [10]

<script type="text/x-mathjax-config">
   MathJax.Hub.Config({  "HTML-CSS": { minScaleAdjust: 125, availableFonts: [] }  });
</script>
</body></html>