### 2019-04-29 STAT0017_ICA2/17052480.Rmd
trans to RMD

---
title: 'STAT0017 ICA2'
author: "Hongwei Peng"
date: "25/04/2019"
output: pdf_document
---

#(a)

```{r message = FALSE}
load("ICA2_data.RData")
library("fUnitRoots")
library(CDVine)
library(fGarch)
library(goftest)
library(KScorrect)
library(stats)
```
```{r}
plot(data$ftse100~as.Date(data$date,"%d/%m/%y"),type="l",xaxt='n',yaxt='n',
     xlab="",ylab="Price",col="blue",main="FTSE100 (prices)",xaxs="i", 
     yaxs="i",ylim=c(3000,8000),cex.main=0.8,cex.lab=1)
axis(2, at = seq(3000,8000,1000), tick=TRUE,cex.axis=0.9)
axis.Date(1, cex.axis=0.9,at=seq(as.Date("1999/02/25"), as.Date("2019/02/28"), "4 years"))

plot(data$sp500~as.Date(data$date,"%d/%m/%y"),type="l",xaxt='n',yaxt='n',
     xlab="",ylab="Price",col="blue",main="S&P500 (prices)",xaxs="i", 
     yaxs="i",ylim=c(600,3000),cex.main=0.8,cex.lab=1)
axis(2, at = seq(600,3000,500), tick=TRUE,cex.axis=0.9)
axis.Date(1, cex.axis=0.9,at=seq(as.Date("1999/02/25"), as.Date("2019/02/28"), "4 years"))

plot(data$sse~as.Date(data$date,"%d/%m/%y"),type="l",xaxt='n',yaxt='n',
     xlab="",ylab="Price",col="blue",main="SSE (prices)",xaxs="i", 
     yaxs="i",ylim=c(1000,6500),cex.main=0.8,cex.lab=1)
axis(2, at = seq(1000,6500,500), tick=TRUE,cex.axis=0.9)
axis.Date(1, cex.axis=0.9,at=seq(as.Date("1999/02/25"), as.Date("2019/02/28"), "4 years"))

unitrootTest(data$ftse100)
unitrootTest(data$sp500)
unitrootTest(data$sse)

ret1<-diff(log(data$ftse100), lag=1,na=remove)
ret2<-diff(log(data$sp500), lag=1,na=remove)
ret3<-diff(log(data$sse), lag=1,na=remove)

plot(ret1~as.Date(data$date[2:length(data$date)],"%d/%m/%y"),type="l",yaxt='n',xaxt='n',
     xlab="",ylab="log-returns",main="FTSE100",xaxs="i", 
     yaxs="i", col="blue",ylim=c(-0.15,0.15),cex.main=0.8,cex.lab=0.8)
axis(2, at = seq(-0.15,0.15,0.1), tick=TRUE,cex.axis=0.7)
axis.Date(1, cex.axis=0.7, at=seq(as.Date("1998/01/04"), as.Date("2019/02/28"), "3 years"))

plot(ret2~as.Date(data$date[2:length(data$date)],"%d/%m/%y"),type="l",yaxt='n',xaxt='n',
     xlab="",ylab="log-returns",main="S&P500",xaxs="i", 
     yaxs="i", col="blue",ylim=c(-0.22,0.2),cex.main=0.8,cex.lab=0.8)
axis(2, at = seq(-0.15,0.15,0.1), tick=TRUE,cex.axis=0.7)
axis.Date(1, cex.axis=0.7, at=seq(as.Date("1998/01/04"), as.Date("2019/02/28"), "3 years"))

plot(ret3~as.Date(data$date[2:length(data$date)],"%d/%m/%y"),type="l",yaxt='n',xaxt='n',
     xlab="",ylab="log-returns",main="S&P500",xaxs="i", 
     yaxs="i", col="blue",ylim=c(-0.22,0.2),cex.main=0.8,cex.lab=0.8)
axis(2, at = seq(-0.15,0.15,0.1), tick=TRUE,cex.axis=0.7)
axis.Date(1, cex.axis=0.7, at=seq(as.Date("1998/01/04"), as.Date("2019/02/28"), "3 years"))

unitrootTest(ret1)
unitrootTest(ret2)
unitrootTest(ret3)

acf(ret1)
acf(ret1^2)

acf(ret2)
acf(ret2^2)

acf(ret3)
acf(ret3^2)

model1=garchFit(formula=~arma(3,0)+garch(1,1),data=ret1,trace=F,cond.dist="sstd")
res1 <- residuals(model1, standardize=TRUE)
acf(res1)
acf(res1^2)
Box.test(res1, lag = 10, type = c("Ljung-Box"), fitdf = 0)
Box.test(res1^2, lag = 10, type = c("Ljung-Box"), fitdf = 0)
shape1<-coef(model1)[9]
skew1<-coef(model1)[8]
u1<-psstd(res1, mean=0, sd=1, nu=shape1, xi=skew1)
hist(u1)

#Kolmogorov-Smirnov test
KStest1<-LcKS(u1, cdf = "punif")
KStest1$p.value
#Anderson-Darling test
ADtest1<-ad.test(u1, null="punif")
ADtest1$p.value

model2=garchFit(formula=~arma(7,0)+garch(1,1),data=ret2,trace=F,cond.dist="sstd")
res2 <- residuals(model2, standardize=TRUE)
acf(res2)
acf(res2^2)
Box.test(res2, lag = 10, type = c("Ljung-Box"), fitdf = 0)
Box.test(res2^2, lag = 10, type = c("Ljung-Box"), fitdf = 0)
shape2<-coef(model2)[13]
skew2<-coef(model2)[12]
u2<-psstd(res2, mean=0, sd=1, nu=shape2, xi=skew2)
#u2<-pnorm(res2, mean=0, sd=1)
#u2<-pt(res2, df=3)
hist(u2)

#Kolmogorov-Smirnov test
KStest2<-LcKS(u2, cdf = "punif")
KStest2$p.value
#Anderson-Darling test
ADtest2<-ad.test(u2, null="punif")
ADtest2$p.value

model3=garchFit(formula=~arma(3,0)+garch(1,1),data=ret3,trace=F,cond.dist="sstd")
res3 <- residuals(model3, standardize=TRUE)
acf(res3)
acf(res3^2)
Box.test(res3, lag = 10, type = c("Ljung-Box"), fitdf = 0)
Box.test(res3^2, lag = 10, type = c("Ljung-Box"), fitdf = 0)
shape3<-coef(model3)[9]
skew3<-coef(model3)[8]
u3<-psstd(res3, mean=0, sd=1, nu=shape3, xi=skew3)
hist(u3)

#Kolmogorov-Smirnov test
KStest3<-LcKS(u3, cdf = "punif")
KStest3$p.value
#Anderson-Darling test
ADtest3<-ad.test(u3, null="punif")
ADtest3$p.value

```

#(b)

#(c)

#(d)