### 2019-07-04 MSc Project/scaled_PLSR_for_corn_data.Rmd
test

---
title: "Scaled PLSR for corn data"
author: "Hongwei Peng"
date: "04/07/2019"
output: pdf_document
---
```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(m5data),                  #input data
                   y=propvals[,1])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 1~m5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 1~m5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(m5data),                  #input data
                   y=propvals[,2])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 2~m5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 2~m5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(m5data),                  #input data
                   y=propvals[,3])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 3~m5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 3~m5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(m5data),                  #input data
                   y=propvals[,4])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 4~m5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 4~m5data Loop=200 scaled")
apply(PlotData,2,mean)
```

###############

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp5data),                  #input data
                   y=propvals[,1])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 1~mp5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 1~mp5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp5data),                  #input data
                   y=propvals[,2])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 2~mp5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 2~mp5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp5data),                  #input data
                   y=propvals[,3])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 3~mp5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 3~mp5data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp5data),                  #input data
                   y=propvals[,4])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 4~mp5data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 4~mp5data Loop=200 scaled")
apply(PlotData,2,mean)
```

#################

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp6data),                  #input data
                   y=propvals[,1])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 1~mp6data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 1~mp6data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp6data),                  #input data
                   y=propvals[,2])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 2~mp6data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 2~mp6data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp6data),                  #input data
                   y=propvals[,3])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 3~mp6data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 3~mp6data Loop=200 scaled")
apply(PlotData,2,mean)
```

```{r echo=FALSE}
library(R.matlab)
library(pls) 
library(lars)
library(ggplot2)
rawdata <- readMat("/Users/hongwei/Downloads/corn.mat")
m5data <-  apply(rawdata$m5spec$data,2,scale)
mp5data <-  apply(rawdata$mp5spec$data,2,scale)
mp6data <-  apply(rawdata$mp6spec$data,2,scale) 
propvals <- apply(rawdata$propvals$data,2,scale)

corn_PLS=function(n){                                #n is the number of calibration
  NV <- 10                                           #number of variables 
  sample <- sample(1:80)                             #set random order; the begin of reset order
  DF <- data.frame(NIR = I(mp6data),                  #input data
                   y=propvals[,4])
  class(DF$NIR) <- "matrix"                          # just to be certain, it was "AsIs"
  #str(DF)                                           #check point
  DF$train <- rep(FALSE, 80)
  DF$train[sample<=n] <- TRUE                        #chose calibration
  corn.pls <- plsr(y ~ NIR, data = DF, ncomp = NV, validation="LOO", jackknife = TRUE, subset = train)
  #summary(corn.pls,what="all")                      #check point
  RMSECV <- RMSEP(corn.pls)$val[1,1,NV]
  #print(RMSECV)                                     #check point
  predict <- predict(corn.pls, ncomp = NV, newdata = DF[!DF$train,])
  RMSEP <- sqrt(sum((predict-DF[!DF$train,]$y)^2)/(80-n))
  #print(RMSEP)                                      #check point
  #plot(R2(corn.pls))                                #check point
  return(cbind(RMSECV,RMSEP))                        #return 1x2matrix
}
#corn_PLS(n)
n <- as.matrix(rep(20:70,200))                        #the number of calibration, rep(a:b,c): from a to b and repeat c. 
PlsResult <- apply(n,1,corn_PLS)                     #loop
PlotData <- as.data.frame(cbind(n,t(PlsResult)))     #combind the results
par(mfrow=c(1,2))
boxplot(V2~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSECV",main="PLS 4~mp6data Loop=200 scaled")
boxplot(V3~V1,data=PlotData,xlab="Number of Calibration", ylab="RMSEP",main="PLS 4~mp6data Loop=200 scaled")
apply(PlotData,2,mean)
```